---
title: "The Singleton Pattern"
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
```

<!-- One line about what the function does -->

***Singleton*** ensures a class only has one instance, and provide a global
point of access to it.

## How It Works

1.  Create only one instance of the **Singleton** class; and
2.  If an instance exists, then serve the same object again.

The main features of **Singleton** are:

-   Ensuring that one and only one object of the class gets created;

-   Providing an access point for an object that is global to the program; and

-   Controlling concurrent access to resources that are shared.

## When to Use It

-   In situations that require exactly one instance of a class, that must be
    accessible to clients from a well-known access point. See the [`Counter`
    example](#example-1).

```{r, error=TRUE}
stop("Singletons can be a problem in multi-threaded applications, especially when they manipulate mutable data.")
```

```{r, message=TRUE}
message("Singletons work well for immutable data, such as reading from some data source, since anything that can’t change isn’t going to run into thread clash problems.")
```

