---
title: "The Value Object Pattern"
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
NA_POSIXct_ <- .POSIXct(NA_real_, tz = "UTC")
is.POSIXct <- function(x) inherits(x, "POSIXct")
```

<!-- One line about what the function does -->

A **Value Object** models a domain concept using natural lingo of the domain
experts, such as "Passenger", "Address", and "Money".

Any **Value Object** is created by a function that receives input, applies some
transformations, and outputs the results in some data structure such as a
vector, a list or a data.frame.

<div>

<mark>**Value Objects** are a vital building block of analytic
applications.</mark>

</div>

## How It Works

In R, a good option for creating a **Value Object** is to follow two
instructions:

-   A **Value Object** is created by a `function`, rather than a class method;
    and
-   A **Value Object** returns a `tibble`, rather than a list or a vector.

In essence, a **Value Object** is a data type, like `integer`, `logical`, `Date`
or `data.frame` data types to name a few. While the built-in data types in R fit
any application, **Value Objects** are domain specific and as such, they fit
only to a specific application. This is because, `integer` is an abstract that
represent whole numbers. This abstract is useful in any application. However, a
**Value Object** represent a high-level abstraction that appears in a particular
domain.

An example of a **Value Object** is the notion of a "Person". Any person in the
world has a name. Needless to say, a person name is spelt by letters, rather
than numbers. A **Value Object** captures these attribute as `tibble` columns
and type checks:

```{r, echo=TRUE}
Person <- function(given = NA_character_, family = NA_character_){
    stopifnot(is.character(given), is.character(family))
    stopifnot(length(given) == length(family))
    
    return(
        tibble::tibble(given = given, family = family)
        %>% tidyr::drop_na(given)
    )
}
```

Instantiating a person **Value Object** is done by calling the `Person`
constructor function:

```{r, echo=TRUE}
person <- Person(given = "Bilbo", family = "Baggins")
```

Getting to know the advantages of a **Value Object**, we should consider the
typical alternative -- constructing a Person by using the `tibble` function
directly:

```{r, echo=TRUE}
person <- tibble::tibble(given = "Bilbo", family = "Baggins")
```

Both implementations return objects with identical content and structure, that
is, their column names, column types and cell values are identical. Then, why
would one prefer using a **Value Object** and its constructor over the direct
alternative?

There are four predominant qualities offered by the **Value Object** pattern
which are not offered by the alternative:

1.  Readability. Each **Value Object** captures a concept belonging to the
    problem domain. Rather than trying to infer what a `tibble` is by looking at
    its low-level details, the **Value Object** constructor descries a context
    on a high-level.

2.  Explicitness. Since the constructor of the **Value Object** is a function,
    its expected input arguments and their type can be detailed in a helper
    file. Moreover, assigning input arguments with default values of specific
    type, such as `NA` (logical NA), `NA_integer_`, `NA_character_`, or
    `NA_Date` (see `lubridate::NA_Date`), expresses clearly the variable types
    of the **Value Object**.

3.  Coherence. The representation of a **Value Object** is concentrated in one
    place -- its constructor. Any change, mainly modifications and extensions,
    applied to the constructor promise the change would propagate to all
    instances of the Value Objects. That means, no structure discrepancies
    between instances that are supposed to represent the same concept.

4.  Safety. The constructor may start with [defensive
    programming](https://en.wikipedia.org/wiki/Defensive_programming) to ensure
    the qualities of its input. One important assertion is type checking. Type
    checking eliminated the risk of implicit type coercing. Another important
    assertion is checking if the lengths of the input arguments meet some
    criteria, say all inputs are of the same length, or more restrictively, all
    inputs are scalars. Having a set of checks makes the code base more robust.
    This is because **Value Objects** are regularly created with the output of
    other functions calls, having a set of checks serves as pseudo-tests of
    these functions output throughout the code.

In addition to these qualities, there are two desirable behaviours which are not
offered by directly calling `tibble`:

1.  [Null Value
    Object]((https://tidylab.github.io/R6P/articles/patterns/NullObject.html)).
    Calling the **Value Object** constructor with no input arguments returns the
    structure of the `tibble` (column names and column types).

2.  Default values for missing input arguments. In this manner, the **Value
    Object** has a well-defined behaviour for a person without a family name,
    such as Madonna and Bono.

In addition to native R data types, a **Value Object** constructor can receive
other **Value Objects** as input arguments. Here are two examples that transmute
Person to other Person-based concepts:

```{r, echo = TRUE, eval = TRUE}
# A Passenger is a Person with a flight booking reference
Passenger <- function(person = Person(), booking_reference = NA_character_){
    stopifnot(all(colnames(person) %in% colnames(Person())))
    stopifnot(is.character(booking_reference))
    
    return(
        person 
        %>% tibble::add_column(booking_reference = booking_reference)
        %>% tidyr::drop_na(booking_reference)
    )
}

person <- Person(given = "Bilbo", family = "Baggins")
passenger <- Passenger(person = person, booking_reference = "B662HR")
print(passenger)
```

```{r, echo = TRUE, eval = TRUE}
# A Diner is a Person that may have dinner reservation
Diner <- function(person = Person(), reservation_time = NA_POSIXct_){
    stopifnot(all(colnames(person) %in% colnames(Person())))
    stopifnot(is.POSIXct(reservation_time))
    
    return(
        person 
        %>% tibble::add_column(reservation_time = reservation_time)
    )
}

person <- Person(given = "Bilbo", family = "Baggins")
timestamp <- as.POSIXct("2021-01-23 18:00:00 NZDT")
diner <- Diner(person = person, reservation_time = timestamp)
print(diner)
```

## When to Use It

-   In situations where domain concepts are more important then the database
    schema. For example, when you are modelling Passengers, your first instinct
    might be to think about the different data sources you'd need for the
    analysis. You may envision "FlightDetails" and "CustomerDetails". Next you
    will define the relationship between them. Instead, let the domain drive the
    design. Create a Passenger **Value Object** with the attributes you must
    have, regardless of any particular database schema.

-   In a function that runs within a specific context. Rather than having an
    input argument called `data` of type `data.frame`, use the appropriate
    **Value Object** name and pass it its constructor.

```{r, echo = TRUE, eval = FALSE}
Audience <- Person

## Without a Value Object
clean_audience_data <- function(data) 
    dplyr::mutate(.data = data, given = stringr::str_to_title(given))

## With a Value Object
clean_audience_data <- function(attendees = Audience()) 
    dplyr::mutate(.data = attendees, given = stringr::str_to_title(given))
```
