---
title: "The Repository Pattern"
bibliography: [../../inst/REFERENCES.bib]
biblio-style: apalike
link-citations: yes
nocite: | 
  @Fowler2002
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes", "_common.R"))
```

<!-- Details -->

```{r, echo=FALSE, results='asis'}
cat(knitr::knit_child('./vignettes/details/_Repository.Rmd', quiet = TRUE))
```

## Implementations

The code of the abstract base class of **Repository** is

```{r}
path_snippet <- usethis::proj_path("R", "object_relational-Repository.R")
snippet <- readLines(path_snippet)
snippet <- snippet |>
  discard_comments() |>
  discard_null() |>
  discard_empty_lines()
```

```{r, code = snippet, eval = FALSE, echo = TRUE}
```

```{r, message=TRUE}
message("By passing the input argument `inherit = Singleton`, the `AbstractRepository` inherits the qualities of the [**Singleton** pattern](https://tidylab.github.io/R6P/articles/patterns/Singleton.html).")
```

The given implementing of `AbstractRepository` requires you to define four
functions:

(1) `initialize` establishes a database connection of some sort;
(2) `add` adds one or more domain objects into the database;
(3) `del` deletes one or more domain objects from the database; and
(4) `get` retrieve one or more domain objects from the database.

<!-- In addition, there are context depended function you may want to ad -->

```{r, message=TRUE}
message("In general, the **Repository** patterns requires at least the `add` and `get` operations. However, you may rename those operations to fit your context. For example, if you use **Repository** to access various tables in a database, `write_table` and `read_table` might be better names.")
```

```{r, warning=TRUE}
warning("It is up to you to devise a policy that defines (A) what to do when the same entity is added to the Repository; and (B) what to do when a query matches no results.")
```

Each **Repository** implementation is project specific. The following
implementation is a **Repository** of car models with their specifications.

From the caller perspective, both implementations behave identically -- they
have the same queries. Nevertheless, under the hood the two implementations
employ different storage approaches.

### Example: Transient Storage Implementation with [`collections`](https://github.com/randy3k/collections)

```{r, message=TRUE}
message("Transient implementations are a temporal solution that is good for testing and rapid prototyping.")
```

1.  Transient implementations contribute to rapid development because:

-   They can be used before you establish/get access to a real database.
-   They are fast to establish in comparison to DBMS

2.  Transient implementations are useful during testing because they are
    independent of the real database (if any), that means:

-   They start as empty storage allowing the programmer to test specific
    behaviour of the caller.
-   They don't make unintended changes in the real database

```{r, error=TRUE}
stop("Transient implementations are not recommended during the production stage. Transient storage is lost when a session is rebooted. You should think about what are the ramifications of losing all the data put into storage.")
```

First, we define the class constructor, `initialize`, to establish a transient
data storage. In this case we use a dictionary from the [`collections`
package](https://randy3k.github.io/collections/reference/dict.html).

Second, we define the `add`, `del` and `get` functions that operate on the
dictionary.

As an optional step, we define the NULL object. In this case, rather then the
reserved word `NULL`, the NULL object is a data.frame with 0 rows and predefined
column.

```{r Repository-Transient-implementation, echo=TRUE, results='markup'}
TransientRepository <- R6::R6Class(
  classname = "Repository", inherit = R6P::AbstractRepository, public = list(
    initialize = function() {
      private$cars <- collections::dict()
    },
    add = function(key, value) {
      private$cars$set(key, value)
      invisible(self)
    },
    del = function(key) {
      private$cars$remove(key)
      invisible(self)
    },
    get = function(key) {
      return(private$cars$get(key, default = private$NULL_car))
    }
  ), private = list(
    NULL_car = cbind(uid = NA_character_, datasets::mtcars)[0, ],
    cars = NULL
  )
)
```

Adding customised operations is also possible via the R6 `set` function. The
following example, adds a query that returns all the objects in the database

```{r Repository-Transient-implementation-addon, echo=TRUE, results='markup'}
TransientRepository$set("public", "get_all_cars", overwrite = TRUE, function() {
  result <- private$cars$values() |> dplyr::bind_rows()
  if (nrow(result) == 0) {
    return(private$NULL_car)
  } else {
    return(result)
  }
})
```

In this example, we use the `mtcars` dataset with a `uid` column that uniquely
identifies the different cars in the **Repository**:

```{r, echo=TRUE, results='markup'}
mtcars <- datasets::mtcars |> tibble::rownames_to_column("uid")
head(mtcars, 2)
```

Here is how the caller uses the `Repository`:

```{r, echo=TRUE, results='markup'}
## Instantiate a repository object
repository <- TransientRepository$new()

## Add two different cars specification to the repository
repository$add(key = "Mazda RX4", value = dplyr::filter(mtcars, uid == "Mazda RX4"))
repository$add(key = "Mazda RX4 Wag", value = dplyr::filter(mtcars, uid == "Mazda RX4 Wag"))

## Get "Mazda RX4" specification
repository$get(key = "Mazda RX4")

## Get all the specifications in the repository
repository$get_all_cars()

## Delete "Mazda RX4" specification
repository$del(key = "Mazda RX4")

## Get "Mazda RX4" specification
repository$get(key = "Mazda RX4")
```

### Example: Persistent Storage Implementation with [`DBI`](https://dbi.r-dbi.org/)

First, we define the class constructor, `initialize`, to establish an SQLite
database.

Second, we define the `add`, `del` and `get` functions that operate on the
dictionary.

As an optional step, we define the NULL object. In this case, rather then the
reserved word `NULL`, the NULL object is a data.frame with 0 rows and predefined
column.

```{r Repository-Persistent-implementation, echo=TRUE}
PersistentRepository <- R6::R6Class(
  classname = "Repository", inherit = AbstractRepository, public = list(
    #' @param immediate (`logical`) Should queries be committed immediately?
    initialize = function(immediate = TRUE) {
      private$immediate <- immediate
      private$conn <- DBI::dbConnect(RSQLite::SQLite(), dbname = ":memory:")
      DBI::dbCreateTable(private$conn, "mtcars", private$NULL_car)
    },
    add = function(key, value) {
      car <- private$NULL_car |> tibble::add_row(value)
      self$del(key = key)
      DBI::dbAppendTable(private$conn, "mtcars", car)
      invisible(self)
    },
    del = function(key) {
      statement <- paste0("DELETE FROM mtcars WHERE uid = '", key, "'")
      DBI::dbExecute(private$conn, statement, immediate = private$immediate)
      invisible(self)
    },
    get = function(key) {
      statement <- paste0("SELECT * FROM mtcars WHERE uid = '", key, "'")
      result <- DBI::dbGetQuery(private$conn, statement)
      if (nrow(result) == 0) {
        return(private$NULL_car)
      } else {
        return(result)
      }
    }
  ), private = list(
    NULL_car = cbind(uid = NA_character_, datasets::mtcars)[0, ],
    immediate = NULL,
    conn = NULL
  )
)
```

Adding customised operations is also possible via the R6 `set` function. The
following example, adds a query that returns all the objects in the database

```{r Repository-Persistent-implementation-addon, echo=TRUE}
PersistentRepository$set("public", "get_all_cars", overwrite = TRUE, function() {
  statement <- "SELECT * FROM mtcars"
  result <- DBI::dbGetQuery(private$conn, statement)
  if (nrow(result) == 0) {
    return(private$NULL_car)
  } else {
    return(result)
  }
})
```

In this example, we use the `mtcars` dataset with a `uid` column that uniquely
identifies the different cars in the **Repository**:

```{r, echo=TRUE, results='markup'}
mtcars <- datasets::mtcars |> tibble::rownames_to_column("uid")
head(mtcars, 2)
```

Here is how the caller uses the `Repository`:

```{r, echo=TRUE, results='markup'}
## Instantiate a repository object
repository <- PersistentRepository$new()

## Add two different cars specification to the repository
repository$add(key = "Mazda RX4", value = dplyr::filter(mtcars, uid == "Mazda RX4"))
repository$add(key = "Mazda RX4 Wag", value = dplyr::filter(mtcars, uid == "Mazda RX4 Wag"))

## Get "Mazda RX4" specification
repository$get(key = "Mazda RX4")

## Get all the specifications in the repository
repository$get_all_cars()

## Delete "Mazda RX4" specification
repository$del(key = "Mazda RX4")

## Get "Mazda RX4" specification
repository$get(key = "Mazda RX4")
```

## Further Reading

[Repository at Martin Fowler
Blog](https://martinfowler.com/eaaCatalog/repository.html)
