---
title: "The Singleton Pattern"
bibliography: [../../inst/REFERENCES.bib]
biblio-style: apalike
nocite: | 
  @Gamma1995 
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
```

```{r child="../details/Singleton.Rmd"}
```

## Example: Counter {#example-1}

In this example we implement a `Counter` that inherits the qualities of
**Singleton**.

```{r Singleton-Counter-implementation, echo=TRUE, results='markup'}
Counter <<- R6::R6Class(inherit = Singleton, public = list(
    count = 0,
    add_1 = function(){self$count = self$count + 1; invisible(self)}
))
```

Whenever we call the constructor on `Counter`, we always get the exact same
instance:

```{r Singleton-instantiation, echo=TRUE, results='markup'}
counter_A <- Counter$new()
counter_B <- Counter$new()

sprintf("counter_A was crated at %s", pryr::address(counter_A))
sprintf("counter_B was crated at %s", pryr::address(counter_B))

identical(counter_A, counter_B)
```

The two objects are equal and located at the same address; thus, they are the
same object.

When we make a change in any of the class instances, the rest of the instances
are changed as well.

```{r Singleton-Counter-instantiation, echo=TRUE, results='markup'}
# How many times has the counter been increased?
counter_A$count

# Add the counter by 1
counter_A$add_1()

# How many times has the counter been increased?
counter_A$count
counter_B$count 
```

## Example: Data Transfer Object (DTO)

In this example, we crate access to a database that only establishes one
connection per session.

Notice how the `initialize` public method first calls the `initialize` of the
**Singleton**, and then appends its own initialisation routine.

```{r Singleton-DTO-implementation, echo=TRUE, results='hold', eval=TRUE}
DTO <<- R6::R6Class(classname = "DTO", inherit = Singleton, public = list(
    con = NULL,
    initialize = function(){
        super$initialize()
        self$establish_connection()
    },
    establish_connection = function(){
        self$con <- DBI::dbConnect(RSQLite::SQLite(), dbname = ":memory:")
        return(self)
    },
    dbSendQuery = function(statement){
        res <- DBI::dbSendQuery(self$con, statement)
        return(res)
    }
))
```

Once a database connection has established during the session, other calls to
`DTO` instantiate objects that use the same database connection.

```{r Singleton-DTO-instantiation, echo=TRUE, results='markup', eval=TRUE}
database_A <- DTO$new()
database_B <- DTO$new()
identical(database_A, database_B)
```

## Further Reading

[Singleton at Wikipedia](https://en.wikipedia.org/wiki/Singleton_pattern)
