---
title: "The Value Object Pattern"
bibliography: [../../inst/REFERENCES.bib]
biblio-style: apalike
nocite: | 
  @Vernon2013 
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes", "_common.R"))
```

<!-- Details -->

```{r, echo=FALSE, results='asis'}
cat(knitr::knit_child('./vignettes/details/_ValueObject.Rmd', quiet = TRUE))
```

<!-- Example -->

## Example: Person and Minister Value Objects

In this example, we are appointing elected officials to random ministries, just
like in real life.

The nomination process comprises three components: input, function, output.

-   The input is the given and family names of a **Person**
-   The function receives a **Person** value object and pairs it with ministry
    titles
-   The output is a **Minister** value object

**Person** -\> appoint_random_ministries

First, we implement the input type. `Person()` is the constructor of the
**Person** value object.

```{r, echo=TRUE}
#' @title Person Value Object Constructor
#' @description A Person encapsulates the information that constitute an individual
#' @param given (`character`) Individual first name.
#' @param family (`character`) Individual last name.
#' @return (`Person`) Person value object.
Person <- function(given = NA_character_, family = NA_character_) {
  stopifnot(is.character(given), is.character(family))
  stopifnot(length(given) == length(family))

  return(
    tibble::tibble(given = given, family = family)
    |> tidyr::drop_na(given)
  )
}

Person()
```

In addition, we implement a predicate that validates whether an arbitrary
`data.frame` is a **Person** value object.

```{r, echo=TRUE}
# Create a test for objects of type Person
# * Extract the column names of Person by using its Null Object (returned by Person())
# * Check that the input argument has all the columns that a Person has
is.Person <- function(x) all(colnames(x) %in% colnames(Person()))
```

Second, we implement the output type. `Minister()` is the constructor of the
**Minister** value object.

```{r, echo=TRUE}
#' @title Minister Value Object Constructor
#' @decription A 'Minister' is a 'Person' with a ministry title.
#' @param Person (`Person`) A member of parliament.
#' @param title (`character`) A string with one or more ministry titles.
#' @return (`Minister`) Minister value object.
Minister <- function(member = Person(), title = NA_character_) {
  stopifnot(is.Person(member), is.character(title))
  stopifnot(nrow(member) == length(title) | all(is.na(title)))

  member |> dplyr::mutate(title = title)
}
```

Third, we write a function that transforms **Person** into **Minister**.

```{r, echo=TRUE}
#' @title Appoint Parliament Members to Ministries
#' @decription
#' Given one or more parliament members
#' When appoint_random_ministries is called
#' Then the parliament members are appointed to an office.
#' @param memeber (`Person`) A Person value object.
#' @return (`Minister`) Minister value object.
appoint_random_ministries <- function(member = Person()) {
  positions <- c(
    "Arts, Culture and Heritage", "Finance", "Corrections",
    "Racing", "Sport and Recreation", "Housing", "Energy and Resources",
    "Education", "Public Service", "Disability Issues", "Environment",
    "Justice", "Immigration", "Defence", "Internal Affairs", "Transport"
  )

  Minister(member = member, title = sample(positions, size = nrow(member)))
}
```

Finally, we pair parliament members with ministries

```{r, echo=TRUE}
# Listing New Zealand elected officials in 2020, we instantiate a Person Object,
# appoint them to random offices and return a Member value object.
set.seed(2020)

parliament_members <- Person(
  given = c("Jacinda", "Grant", "Kelvin", "Megan", "Chris", "Carmel"),
  family = c("Ardern", "Robertson", "Davis", "Woods", "Hipkins", "Sepuloni")
)

parliament_members

appoint_random_ministries(member = parliament_members)
```

## Further Reading

[Value Object on Wikipedia](https://en.wikipedia.org/wiki/Value_object)
